Experiment 3

### 2024-03-17 Analysis/e3.qmd
add lightbox
e2 and e3 stats. re-render

---
title: "HTW E3"
categories: [Analyses, R, Bayesian]
code-fold: true
cold-tools: true
---

```{r}
pacman::p_load(dplyr,purrr,tidyr,tibble,ggplot2,
  brms,tidybayes, rstanarm,emmeans,broom,bayestestR,
  stringr, here,conflicted, patchwork, knitr,kableExtra)

walk(c("brms","dplyr","bayestestR"), conflict_prefer_all, quiet = TRUE)
walk(c("Display_Functions","org_functions"), ~ source(here::here(paste0("Functions/", .x, ".R"))))

e3 <- readRDS(here("data/e3_08-04-23.rds")) |> 
    mutate(trainCon=case_when(
    bandOrder=="Original" ~ "800",
    bandOrder=="Reverse" ~ "600",
    TRUE ~ NA_character_
    ), trainCon=as.numeric(trainCon))

e3Sbjs <- e3 |> group_by(id,condit,bandOrder) |> summarise(n=n())
testE3 <- e3 |> filter(expMode2 == "Test")

@@

@@

nbins=6
trainE3 <-  e3 |> filter(expMode2=="Train") |> group_by(id,condit,bandOrder, vb) |> 
    mutate(Trial_Bin = cut( gt.train, breaks = seq(1, max(gt.train),length.out=nbins+1),include.lowest = TRUE, labels=FALSE)) 
# bayesian comparison of condits in training
trainE3_max <- trainE3 |> filter(Trial_Bin == nbins, bandInt==trainCon)

@@

#bayestestR::describe_posterior(bayes_comp)

mtr3 <- as.data.frame(describe_posterior(bmm_e3_train, centrality = "Mean"))[, c(1,2,4,5,6)]
colnames(mtr3) <- c("Term", "Estimate","95% CrI Lower", "95% CrI Upper", "pd")

mtr3 |> mutate(across(where(is.numeric), \(x) round(x, 2))) |>
  tibble::remove_rownames() |> 
  mutate(Term = stringr::str_remove(Term, "b_")) |>
  kable(escape=F,booktabs=T) %>% kable_styling(font_size=8)

@@

```{r}
#| label: fig-e3-train-dev
#| fig-cap: E3. Deviations from target band during testing without feedback stage.

trainE3 |> ggplot(aes(x = Trial_Bin, y = dist, color = condit)) +
    stat_summary(geom = "line", fun = mean) +
    stat_summary(geom = "errorbar", fun.data = mean_se, width = .4, alpha = .7) +
    ggh4x::facet_nested_wrap(~bandOrder*vb,ncol=3)+
    scale_x_continuous(breaks = seq(1, nbins + 1)) +
    theme(legend.title=element_blank()) + 
    labs(y = "Deviation", x="Training Block")

@@

@@

#bayestestR::describe_posterior(bmtd)

mted3 <- as.data.frame(describe_posterior(bmtd3, centrality = "Mean"))[, c(1,2,4,5,6)]
colnames(mted3) <- c("Term", "Estimate","95% CrI Lower", "95% CrI Upper", "pd")

mted3 |> mutate(across(where(is.numeric), \(x) round(x, 2))) |>
  tibble::remove_rownames() |> 
  mutate(Term = stringr::str_remove(Term, "b_")) |>
  kable(booktabs=TRUE)

ce_bmtd3 <- plot(conditional_effects(bmtd3),points=FALSE,plot=FALSE)
wrap_plots(ce_bmtd3)

@@

```{r}
#| label: fig-e3-test-dev
#| fig-cap: E3. Deviations from target band during testing without feedback stage. 
pe3td <- testE3 |>  ggplot(aes(x = vb, y = dist,fill=condit)) +
    stat_summary(geom = "bar", position=position_dodge(), fun = mean) +
    stat_summary(geom = "errorbar", position=position_dodge(.9), fun.data = mean_se, width = .4, alpha = .7) + 
    facet_wrap(~bandOrder,ncol=1) +
  theme(legend.title=element_blank(),axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5)) +
  labs(x="Band", y="Deviation From Target")

@@

}

pe3ce <- bmtd3 |> emmeans( ~condit *bandOrder*bandType) |>
  gather_emmeans_draws() |>
 condEffects(bandType) + labs(y="Absolute Deviation From Band", x="Band Type") + 
 facet_wrap(~bandOrder,ncol=1)

pe3td + pe3ce

@@

@@

e3_vxBMM <- brm(vx ~ condit * bandOrder * bandInt + (1 + bandInt|id),
                        data=test,file=paste0(here::here("data/model_cache", "e3_testVxBand_RF_5k")),
                        iter=5000,chains=4,silent=0,
                        control=list(adapt_delta=0.94, max_treedepth=13))
GetModelStats(e3_vxBMM ) |> kable(escape=F,booktabs=T, caption="Fit to all 6 bands")

cd3 <- get_coef_details(e3_vxBMM, "conditVaried")
sc3 <- get_coef_details(e3_vxBMM, "bandInt")
intCoef3 <- get_coef_details(e3_vxBMM, "conditVaried:bandInt")

```{r}
#| label: fig-e3-bmm-vx
#| fig-cap: Conditional effect of training condition and Band. Ribbons indicate 95% HDI. The steepness of the lines serves as an indicator of how well participants discriminated between velocity bands. 
#| fig-height: 19
#| fig-width: 12

pe3tv <- testE3 %>% group_by(id,vb,condit,bandOrder) |> plot_distByCondit() + ggh4x::facet_nested_wrap(bandOrder~condit,scale="free_x")

@@

pid_slopes3 <- re |>  
    mutate(id=reorder(id,slope)) |>
  ggplot(aes(y=id, x=slope,fill=condit,color=condit)) + 
    geom_pointrange(aes(xmin=Q2.5+adjust, xmax=Q97.5+adjust)) + 
    theme(legend.title=element_blank()) + 
    labs(x="Estimated Slope", y="Participant")  + 
    ggh4x::facet_nested_wrap(bandOrder~condit,axes="all",scales="free_y")

pe3tv / (pe3vce + pid_den3 + pid_slopes3)

@@

>>>>

---
title: "Experiment 3"
categories: [Analyses, R, Bayesian]
toc: false
format-links: false
page-layout: full
lightbox: true
code-fold: true
cold-tools: true
cache: true
keep-md: true
format:
  html: default
  hugo-md:
    echo: false
    html-math-method: mathjax
    output-file: e3-hugo.md
  gfm:
    echo: true
    output-file: e3-gfm.md
---

```{r}
pacman::p_load(dplyr,purrr,tidyr,tibble,ggplot2,
  brms,tidybayes, rstanarm,emmeans,broom,bayestestR,
  stringr, here,conflicted, patchwork, knitr,kableExtra)
walk(c("brms","dplyr","bayestestR"), conflict_prefer_all, quiet = TRUE)
walk(c("Display_Functions","org_functions"), ~ source(here::here(paste0("Functions/", .x, ".R"))))
e3 <- readRDS(here("data/e3_08-04-23.rds")) |> 
    mutate(trainCon=case_when(
    bandOrder=="Original" ~ "800",
    bandOrder=="Reverse" ~ "600",
    TRUE ~ NA_character_
    ), trainCon=as.numeric(trainCon)) 
e3Sbjs <- e3 |> group_by(id,condit,bandOrder) |> summarise(n=n())
testE3 <- e3 |> filter(expMode2 == "Test")

nbins=5
trainE3 <-  e3 |> filter(expMode2=="Train") |> group_by(id,condit,bandOrder, vb) |> 
    mutate(Trial_Bin = cut( gt.train, breaks = seq(1, max(gt.train),length.out=nbins+1),include.lowest = TRUE, labels=FALSE)) 
# bayesian comparison of condits in training
trainE3_max <- trainE3 |> filter(Trial_Bin == nbins, bandInt==trainCon)

@@

The major manipulation adjustment of experiment 3 is for participants to receive ordinal feedback during training, in contrast to the continuous feedback of the earlier experiments. Ordinal feedback informs participants whether a throw was too soft, too hard, or fell within the target velocity range. Experiment 3 participants were randomly assigned to both a training condition (Constant vs. Varied) and a Band Order condition (original order used in Experiment 1, or the Reverse order of Experiment 2).

@@

@@

# mtr3 <- as.data.frame(describe_posterior(bmm_e3_train, centrality = "Mean"))[, c(1,2,4,5,6)]
# colnames(mtr3) <- c("Term", "Estimate","95% CrI Lower", "95% CrI Upper", "pd")
# mtr3 |> mutate(across(where(is.numeric), \(x) round(x, 2))) |>
#   tibble::remove_rownames() |> 
#   mutate(Term = stringr::str_remove(Term, "b_")) |>
#   kable(escape=F,booktabs=T)

@@

::: {#tbl-e3-train-dist}

| Term                          | Estimate | 95% CrI Lower | 95% CrI Upper |   pd |
|:------------------------------|---------:|--------------:|--------------:|-----:|
| Intercept                     |   121.86 |        109.24 |        134.60 | 1.00 |
| conditVaried                  |    64.93 |         36.99 |         90.80 | 1.00 |
| bandOrderReverse              |     1.11 |        -16.02 |         18.16 | 0.55 |
| conditVaried:bandOrderReverse |   -77.02 |       -114.16 |        -39.61 | 1.00 |

: **Experiment 3 - End of training performance**. The Intercept represents the average of the baseline (constant condition), and the conditVaried coefficient reflects the difference between the constant and varied groups. A larger positive coefficient indicates a greater deviation (lower accuracy) for the varied group.  {.sm}

:::

```{r}
#| label: fig-e3-train-dev
#| fig-cap: E3. Deviations from target band during testing without feedback stage. 
#| eval: FALSE
p1 <- trainE3 |> ggplot(aes(x = Trial_Bin, y = dist, color = condit)) +
    stat_summary(geom = "line", fun = mean) +
    stat_summary(geom = "errorbar", fun.data = mean_se, width = .4, alpha = .7) +
    ggh4x::facet_nested_wrap(~bandOrder*vb,ncol=3)+
    scale_x_continuous(breaks = seq(1, nbins + 1)) +
    theme(legend.title=element_blank()) + 
    labs(y = "Deviation", x="Training Block")

#ggsave(here("Assets/figs/e3_train_deviation.png"), p1, width = 9, height = 8,bg="white")

@@

::: {#fig-e3-train-dev}

![](../Assets/figs/e3_train_deviation.png)

E3. Deviations from target band across training blocks. 
:::

@@

# mted3 <- as.data.frame(describe_posterior(bmtd3, centrality = "Mean"))[, c(1,2,4,5,6)]
# colnames(mted3) <- c("Term", "Estimate","95% CrI Lower", "95% CrI Upper", "pd")
# mted3 |> mutate(across(where(is.numeric), \(x) round(x, 2))) |>
#   tibble::remove_rownames() |> 
#   mutate(Term = stringr::str_remove(Term, "b_")) |>
#   kable(booktabs=TRUE)

ce_bmtd3 <- plot(conditional_effects(bmtd3),points=FALSE,plot=FALSE)
#wrap_plots(ce_bmtd3)

#ggsave(here::here("Assets/figs", "e3_cond_effects_dist.png"), wrap_plots(ce_bmtd3), width=11, height=11, bg="white")

@@

::: {#tbl-e3-bmm-dist}
| Term                                                | Estimate | 95% CrI Lower | 95% CrI Upper |   pd |
|:----------------------------------------------------|---------:|--------------:|--------------:|-----:|
| Intercept                                           |   288.65 |        199.45 |        374.07 | 1.00 |
| conditVaried                                        |   -40.19 |       -104.68 |         23.13 | 0.89 |
| bandTypeExtrapolation                               |   -23.35 |        -57.28 |         10.35 | 0.92 |
| bandOrderReverse                                    |   -73.72 |       -136.69 |        -11.07 | 0.99 |
| conditVaried:bandTypeExtrapolation                  |    52.66 |         14.16 |         90.23 | 1.00 |
| conditVaried:bandOrderReverse                       |   -37.48 |       -123.28 |         49.37 | 0.80 |
| bandTypeExtrapolation:bandOrderReverse              |    80.69 |         30.01 |        130.93 | 1.00 |
| conditVaried:bandTypeExtrapolation:bandOrderReverse |    30.42 |        -21.00 |         81.65 | 0.87 |

: **Experiment 3 testing accuracy**. Main effects of condition and band type (training vs. extrapolation), and the interaction between the two factors. Larger coefficient estimates indicate larger deviations from the baselines (constant & trained bands) - and a positive interaction coefficient indicates disproporionate deviation for the varied condition on the extrapolation bands {.striped .hover .sm}
:::

::: {#fig-e3-test-condEffect}

![](../Assets/figs/e3_cond_effects_dist.png)

E3. A) Deviations from target band during testing without feedback stage. B) Estimated marginal means for the interaction between training condition and band type. Error bars represent 95% confidence intervals.
:::

```{r}
#| label: fig-e3-test-dev
#| fig-cap: E3. Deviations from target band during testing without feedback stage. 
#| eval: FALSE

pe3td <- testE3 |>  ggplot(aes(x = vb, y = dist,fill=condit)) +
    stat_summary(geom = "bar", position=position_dodge(), fun = mean) +
    stat_summary(geom = "errorbar", position=position_dodge(.9), fun.data = mean_se, width = .4, alpha = .7) + 
    facet_wrap(~bandOrder,ncol=1) +
  theme(legend.title=element_blank(),axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5)) +
  labs(x="Band", y="Deviation From Target")

@@

}
pe3ce <- bmtd3 |> emmeans( ~condit *bandOrder*bandType) |>
  gather_emmeans_draws() |>
 condEffects(bandType) + labs(y="Absolute Deviation From Band", x="Band Type") + 
 facet_wrap(~bandOrder,ncol=1)

p2 <- pe3td + pe3ce + plot_annotation(tag_levels= 'A')
#ggsave(here::here("Assets/figs", "e3_test-dev.png"), p2, width=9, height=8, bg="white")

@@

::: {#fig-e3-test-dev}

![](../Assets/figs/e3_test-dev.png)

E3. A) Deviations from target band during testing without feedback stage. B) Estimated marginal means for the interaction between training condition and band type. Error bars represent 95% confidence intervals.
:::

@@

e3_vxBMM <- brm(vx ~ condit * bandOrder * bandInt + (1 + bandInt|id),
                        data=test,file=paste0(here::here("data/model_cache", "e3_testVxBand_RF_5k")),
                        iter=5000,chains=4,silent=0,
                        control=list(adapt_delta=0.94, max_treedepth=13))

# m1 <- as.data.frame(describe_posterior(e3_vxBMM, centrality = "Mean"))
# m2 <- fixef(e3_vxBMM)
# mp3 <- m1[, c(1,2,4,5,6)]
# colnames(mp3) <- c("Term", "Estimate","95% CrI Lower", "95% CrI Upper", "pd")                       
# mp3 |> mutate(across(where(is.numeric), \(x) round(x, 2))) |>
#   tibble::remove_rownames() |> 
#   mutate(Term = stringr::str_replace_all(Term, "b_bandInt", "Band")) |>
#   kable(escape=F,booktabs=T)

cd1 <- get_coef_details(e3_vxBMM, "conditVaried")
sc1 <- get_coef_details(e3_vxBMM, "bandInt")
intCoef1 <- get_coef_details(e3_vxBMM, "conditVaried:bandInt")
```

::: {#tbl-e3-bmm-vx}

| Term                                    | Estimate | 95% CrI Lower | 95% CrI Upper |   pd |
|:----------------------------------------|---------:|--------------:|--------------:|-----:|
| b_Intercept                             |   601.83 |        504.75 |        699.42 | 1.00 |
| b_conditVaried                          |    12.18 |       -134.94 |        162.78 | 0.56 |
| b_bandOrderReverse                      |    13.03 |       -123.89 |        144.67 | 0.58 |
| Band                                    |     0.49 |          0.36 |          0.62 | 1.00 |
| b_conditVaried:bandOrderReverse         |  -338.15 |       -541.44 |       -132.58 | 1.00 |
| b_conditVaried:bandInt                  |    -0.04 |         -0.23 |          0.15 | 0.67 |
| b_bandOrderReverse:bandInt              |    -0.10 |         -0.27 |          0.08 | 0.86 |
| b_conditVaried:bandOrderReverse:bandInt |     0.42 |          0.17 |          0.70 | 1.00 |

Experiment 3. Bayesian Mixed Model Predicting Vx as a function of condition (Constant vs. Varied) and Velocity Band {.striped .hover .sm}
:::

```{r}
#| label: fig-e3-bmm-vx
#| fig-cap: Conditional effect of training condition and Band. Ribbons indicate 95% HDI. The steepness of the lines serves as an indicator of how well participants discriminated between velocity bands. 
#| fig-height: 19
#| fig-width: 12
#| eval: FALSE
pe3tv <- testE3 %>% group_by(id,vb,condit,bandOrder) |> plot_distByCondit() + ggh4x::facet_nested_wrap(bandOrder~condit,scale="free_x")

@@

pid_slopes3 <- re |>  
    mutate(id=reorder(id,slope)) |>
  ggplot(aes(y=id, x=slope,fill=condit,color=condit)) + 
    geom_pointrange(aes(xmin=Q2.5+adjust, xmax=Q97.5+adjust)) + 
    theme(legend.title=element_blank(), 
      axis.text.y = element_text(size=6) ) + 
    labs(x="Estimated Slope", y="Participant")  + 
    ggh4x::facet_nested_wrap(bandOrder~condit,axes="all",scales="free_y")

p3 <- pe3tv / (pe3vce + pid_den3 + pid_slopes3) + plot_annotation(tag_levels= 'A')
#ggsave(here::here("Assets/figs", "e3_test-vx.png"), p3,width=11,height=13, bg="white",dpi=800)
#p3

@@

{{< pagebreak >}}

::: {#fig-e3-test-vx}

![](../Assets/figs/e3_test-vx.png)

Experiment 2. Conditional effect of training condition and Band. Ribbons indicate 95% HDI. The steepness of the lines serves as an indicator of how well participants discriminated between velocity bands.

:::
